---
title: "Decennial DP Evaluation"
format: 
  html:
    embed-resources: true
    toc: true
editor_options: 
  chunk_output_type: console
---

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(here)
library(urbnthemes)
library(tigris)

source(here("R", "metrics_config.R"))
source(here("R", "metric_all.R"))

set_urbn_defaults(style = "print")
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)
```

# Prep

## Summarize Metrics

```{r}
summarize_results <- function(data) {
  
  data |>
    metric_all(n = n, n_noisy = n_noisy, threshold = 0.1, drop_zeros = TRUE) |>
    ungroup()
  
}
```

## Identify Summary Level

```{r}
methods_summary_level <- c("method", "run", "epsilon", "opt_in")
race_state_summary_level <- c("race_simple", "state", "method", "run", "epsilon", "opt_in")
hisp_state_summary_level <- c("hisp_simple", "state", "method", "run", "epsilon", "opt_in")
mult_race_state_summary_level <- c("white_multiplier", "race_simple", "state", "method", "run", "epsilon", "opt_in")
```

## Plotting Functions

```{r}
boxplot_methods <- function(data, metric, ylab) {
  
  data %>%
    ggplot(aes(fill = method, 
               y = {{metric}}, 
               x = as.factor(opt_in))) +
      geom_boxplot() +
      geom_hline(yintercept = 0, color = "black", size = 0.05) +
      labs(x = "Opt-in rate",
           y = ylab) +
      scale_fill_discrete(labels = c("global" = "Central", 
                                     "grr" = "Local")) +
      facet_grid(epsilon ~ ., scales = "free_y", 
                 labeller = label_bquote(epsilon == .(epsilon)))
  
}

boxplot_race_state <- function(data, mech, state, metric, ylab, eps) {
  
  state_fips <- parse_number(lookup_code(state))
  
  data %>%
    filter(state == state_fips & epsilon == eps & method == mech) %>%
    ggplot(aes(y = {{metric}}, 
               x = as.factor(opt_in))) +
      geom_boxplot() +
      geom_hline(yintercept = 0, color = "black", size = 0.05) +
      labs(x = "Opt-in rate",
           y = ylab) +
      facet_grid(race_simple ~ .)
  
}

boxplot_hisp_state <- function(data, mech, state, metric, ylab, eps) {
  
  state_fips <- parse_number(lookup_code(state))
  
  data %>%
    filter(state == state_fips & epsilon == eps & method == mech) %>%
    ggplot(aes(y = {{metric}}, 
               x = as.factor(opt_in))) +
      geom_boxplot() +
      geom_hline(yintercept = 0, color = "black", size = 0.05) +
      labs(x = "Opt-in rate",
           y = ylab) +
    facet_grid(hisp_simple ~ .)

}

boxplot_mult_race_state <- function(data, mech, state, metric, eps, optin) {
  
  state_fips <- parse_number(lookup_code(state))
  
  data %>%
    filter(method == mech & 
           epsilon == eps & 
           state == state_fips & 
           opt_in == optin) %>%
    ggplot(aes(fill = race_simple, 
               y = {{metric}}, 
               x = as.factor(white_multiplier))) +
      geom_boxplot() +
      geom_hline(yintercept = 0, color = "black", size = 0.05) +
      labs(x = "White Multiplier")

}
```

# Decennial 001 No Mult Results

## Prep Data

```{r}
results_nomult001 <- readRDS(here("data", "results", "decennial-results-nomult-001.rds"))
```

```{r}
methods_summary_nomult001 <- results_nomult001 %>%
  group_by(across(all_of(methods_summary_level))) %>%
  summarize_results()

race_state_summary_nomult001 <- results_nomult001 %>%
  group_by(across(all_of(race_state_summary_level))) %>%
  summarize_results()
```

## Results

### Methods Overall

#### Bias

```{r}
methods_summary_nomult001 %>%
  boxplot_methods(mean_pct_error, "Mean percent error")
```

#### Accuracy

```{r}
methods_summary_nomult001 %>%
  boxplot_methods(mean_abs_pct_error, "Mean absolute percent error")
```

### Methods by Race & State

#### Bias

```{r}
race_state_summary_nomult001 %>%
  boxplot_race_state(mech = "grr",
                     state = "DC",
                     metric = mean_pct_error,
                     ylab = "Mean percent error",
                     eps = 1)
```

```{r}
race_state_summary_nomult001 %>%
  boxplot_race_state(mech = "grr",
                     state = "IA",
                     metric = mean_pct_error,
                     ylab = "Mean percent error",
                     eps = 1)
```

#### Accuracy

```{r}
race_state_summary_nomult001 %>%
  boxplot_race_state(mech = "grr",
                     state = "DC",
                     metric = mean_abs_pct_error,
                     ylab = "Mean absolute percent error",
                     eps = 1)
ggsave(here("analysis", "figures", "dec-race-accuracy-dc.png"))
```

```{r}
race_state_summary_nomult001 %>%
  boxplot_race_state(mech = "grr",
                     state = "IA",
                     metric = mean_abs_pct_error,
                     ylab = "Mean absolute percent error",
                     eps = 1)
ggsave(here("analysis", "figures", "dec-race-accuracy-ia.png"))
```

```{r}
race_state_summary_nomult001 %>%
  filter(epsilon == 1, method == "grr", race_simple == "Hispanic") %>%
  group_by(across(all_of(c("state", "opt_in")))) %>%
  summarize(median_mape = median(mean_abs_pct_error))
```

# Decennial 002 No Mult Results

## Prep Data

```{r}
results_nomult002 <- readRDS(here("data", "results", "decennial-results-nomult-002.rds"))
```

```{r}
methods_summary_nomult002 <- results_nomult002 %>%
  group_by(across(all_of(methods_summary_level))) %>%
  summarize_results()

hisp_state_summary_nomult002 <- results_nomult002 %>%
  group_by(across(all_of(hisp_state_summary_level))) %>%
  summarize_results()
```

## Results

### Methods Overall

#### Bias

```{r}
methods_summary_nomult002 %>%
  boxplot_methods(mean_pct_error, "Mean percent error")
ggsave(here("analysis", "figures", "dec-methods-bias.png"))
```

#### Accuracy

```{r}
methods_summary_nomult002 %>%
  boxplot_methods(mean_abs_pct_error, "Mean absolute percent error")
ggsave(here("analysis", "figures", "dec-methods-accuracy.png"))
```

### Methods by Hisp & State

#### Bias

```{r}
hisp_state_summary_nomult002 %>%
  boxplot_hisp_state(state = "DC",
                     mech = "grr",
                     metric = mean_pct_error,
                     ylab = "Mean percent error",
                     eps = 1)
```

```{r}
hisp_state_summary_nomult002 %>%
  boxplot_hisp_state(state = "IA",
                     mech = "grr",
                     metric = mean_pct_error,
                     ylab = "Mean percent error",
                     eps = 1)
```

#### Accuracy

```{r}
hisp_state_summary_nomult002 %>%
  boxplot_hisp_state(state = "DC",
                     mech = "grr",
                     metric = mean_abs_pct_error,
                     ylab = "Mean absolute percent error",
                     eps = 1)
```

```{r}
hisp_state_summary_nomult002 %>%
  boxplot_hisp_state(state = "IA",
                     mech = "grr",
                     metric = mean_abs_pct_error,
                     ylab = "Mean absolute percent error",
                     eps = 1)
```

# Decennial 001 Mult Results

## Prep Data

```{r}
results_mult001 <- readRDS(here("data", "results", "decennial-results-mult-001.rds"))
```

```{r}
mult_race_state_summary_mult001 <- results_mult001 %>%
  group_by(across(all_of(mult_race_state_summary_level))) %>%
  summarize_results()
```

## Results

### Methods by Hisp & State

#### Accuracy

```{r}
boxplot_mult_race_state(data = mult_race_state_summary_mult001,
                        mech = "grr",
                        state = "IA",
                        metric = mean_abs_pct_error,
                        eps = 1,
                        optin = 0.5)
```

```{r}
boxplot_mult_race_state(data = mult_race_state_summary_mult001,
                        mech = "grr",
                        state = "DC",
                        metric = mean_abs_pct_error,
                        eps = 1,
                        optin = 0.5)
```

