---
title: "Synthetic ACS Evaluation"
format: 
  html:
    embed-resources: true
    toc: true
    df-print: paged
    fig-width: 6.5
editor: source
execute:
  echo: true
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
knitr:
  opts_chunk:
    fig.path: figures/
---

```{r}
#| label: load-packages-and-function

library(tidyverse)
library(furrr)
library(here)
library(tidysynthesis)
library(syntheval)
library(tictoc)
library(patchwork)
library(urbnthemes)

set_urbn_defaults(style = "print")

source(here("R", "prediction_test.R"))
source(here("R", "kmarginals.R"))
source(here("R", "proportions.R"))
source(here("R", "discriminator_auc.R"))
source(here("R", "regression_ci_overlap.R"))
source(here("R", "membership_inference_test.R"))

source(here("R", "summarize_results.R"))

theme_urbn_print <- theme_update(strip.background = ggplot2::element_rect(fill = "#ffffff", colour = NA))


```

## Load ACS

```{r}
#| label: load-data
gsds <- readRDS(here("data", "starting", "acs-gsds.rds")) |>
  select(-serial)

holdout <- readRDS(here("data", "starting", "acs-holdout.rds")) |>
  select(-serial)

set.seed(20230822)
holdout_frac <- holdout  |>
  slice_sample(prop = 0.1)

syntheses <- list.files(here::here("data", "results", "acs"), full.names = TRUE)

syntheses <- syntheses[str_detect(syntheses, pattern = "synth-acs_")]

syntheses <- syntheses |>
  map(readRDS)

# syntheses <- syntheses[1:2]

# plan(multisession, workers = 4)
# message("Number of parallel workers: ", nbrOfWorkers())

if (!file.exists(here("data", "results", "acs-metrics.rds"))) {

  set.seed(20230818)
  tic()
  results <-
    map(
      .x = syntheses,
      .f = ~summarize_results(postsynth = .x, data = gsds, holdout = holdout_frac),
      .progress = TRUE
    )
  toc()
  
  saveRDS(results, here("data", "results", "acs-metrics.rds"))
  
} else {
  
  results <- readRDS(here("data", "results", "acs-metrics.rds"))
  
}



specs <- read_csv(here::here("data", "results", "specs.csv")) |>
  mutate(
    opt_in = factor(opt_in),
    white_multiplier = factor(white_multiplier)
  )

```

## Univariate Utility

### Proportions

```{r}
#| label: proportions
#| fig-height: 4

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 32)),
  results |>
    map("proportions") |>
    map_dfr(.f = ~.x |> 
              pivot_wider(id_cols = c(variable, class), names_from = source, values_from = prop)
    ) |>
    mutate(abs_error = abs(original - synthetic))
) |>
  mutate(
    variable = case_match(
      variable,
      "empstat" ~ "Employment",
      "hcovany" ~ "Health insurance\ncoverage",
      "hispan" ~ "Hispanicity",
      "labforce" ~ "Labor force\nparticipation",
      "marst" ~ "Marital status",
      "race" ~ "Race",
      "sex" ~ "Sex",
      "statefip" ~ "State"
    )
  ) |>
  filter(white_multiplier == "1") |>
  ggplot(aes(opt_in, abs_error)) +
  geom_point(alpha = 0.4) +
  facet_wrap(~ variable, nrow = 2) +
  labs(
    x = "Opt-In Rate",
    y = "Absolute Error In Proportion"
  ) +
  scatter_grid()

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 97)),
  results |>
    map("proportions_race_eth") |>
    map_dfr(.f = ~.x |> 
              pivot_wider(id_cols = c(variable, class, race_eth), names_from = source, values_from = prop)
    ) |>
    mutate(error = original - synthetic)
) |>
  filter(white_multiplier == 1) |>
  filter(!variable %in% c("hispan", "race")) |>
  mutate(
    variable = case_match(
      variable,
      "empstat" ~ "Employment",
      "hcovany" ~ "Health insurance coverage",
      "labforce" ~ "Labor force participation",
      "marst" ~ "Marital status",
      "sex" ~ "Sex",
      "statefip" ~ "State"
    )
  ) |>
  ggplot(aes(opt_in, error, color = race_eth)) +
  geom_point(alpha = 0.4) +
  facet_wrap(~ variable, nrow = 2) +
  labs(
    x = "Opt-In Rate",
    y = "Error In Proportion"
  ) +
  scatter_grid()

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 97)),
  results |>
    map("proportions_race_eth") |>
    map_dfr(.f = ~.x |> 
              pivot_wider(id_cols = c(variable, class, race_eth), names_from = source, values_from = prop)
    ) |>
    mutate(abs_error = abs(original - synthetic))
) |>
  filter(white_multiplier == 1) |>
  filter(!variable %in% c("hispan", "race")) |>
  mutate(
    variable = case_match(
      variable,
      "empstat" ~ "Employment",
      "hcovany" ~ "Health insurance coverage",
      "labforce" ~ "Labor force participation",
      "marst" ~ "Marital status",
      "sex" ~ "Sex",
      "statefip" ~ "State"
    )
  ) |>  
  ggplot(aes(opt_in, abs_error, color = race_eth)) +
  geom_point(alpha = 0.4) +
  facet_wrap(~ variable, nrow = 2) +
  labs(
    x = "Opt-In Rate",
    y = "Absolute Error In Proportion"
  ) +
  scatter_grid()

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 97)),
  results |>
    map("proportions_race_eth") |>
    map_dfr(.f = ~.x |> 
              pivot_wider(id_cols = c(variable, class, race_eth), names_from = source, values_from = prop)
    ) |>
    mutate(error = original - synthetic)
) |>
  filter(id == "011", race_eth == "Black", variable == "marst")

```

### Proportions For Three Variables

```{r}
#| label: proportions-special
#| fig-height: 4

bind_cols(
  specs |>
    slice(rep(1:n(), each = 12)),
  results |>
    map("proportions_race_eth") |>
    map_dfr(.f = ~.x |> 
              pivot_wider(id_cols = c(variable, class, race_eth), names_from = source, values_from = prop)) |>
    filter(class %in% c("Never married/single", "Employed", "With health insurance coverage"))
) |>
  filter(white_multiplier == 1) |>
  mutate(diff = original - synthetic) |>
  group_by(opt_in, race_eth, class) |>
  summarize(median_diff = median(diff)) |>
  ggplot(aes(opt_in, median_diff, color = race_eth)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(group = race_eth), alpha = 0.5) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ class) +
  labs(
    x = "Opt-In Rate",
    y = "Median Errors in Estimated Proportions"
  )

```

### Proportion with Wages

```{r}
#| label: proportion-with-wages

bind_cols(
  specs,
  results |>
    map("moments") |>
    map_dfr(.f = ~.x |> filter(variable == "incwage", statistic == "have"))
) |>
  ggplot(aes(opt_in, proportion_difference, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    y = "Error in Proportion with Salaries and Wages",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 9)),
  results |>
    map("moments_race") |>
    map_dfr(.f = ~.x |> filter(variable == "incwage", statistic == "have"))
) |>
  ggplot(aes(opt_in, proportion_difference, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0) +
  facet_wrap(~race) +
  labs(
    x = "Opt-In Rate",
    y = "Error in Proportion with Salaries and Wages",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 4)),
  results |>
    map("moments_race_eth") |>
    map_dfr(.f = ~.x |> filter(variable == "incwage", statistic == "have"))
) |>
  ggplot(aes(opt_in, proportion_difference, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0) +
  facet_wrap(~ race_eth) +
  labs(
    x = "Opt-In Rate",
    y = "Error in Proportion with Salaries and Wages",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

```

### Mean Income

```{r}
#| label: mean-income

bind_cols(
  specs,
  results |>
    map("moments") |>
    map_dfr(.f = ~.x |> filter(variable == "incwage", statistic == "mean"))
) |>
  ggplot(aes(opt_in, proportion_difference, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    y = "Proportion Error in Mean Wage and Salary Income",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

bind_cols(
  specs |> 
    slice(rep(1:n(), each = 9)),
  results |>
    map("moments_race") |>
    map_dfr(.f = ~.x |> filter(variable == "incwage", statistic == "mean"))
) |>
  ggplot(aes(opt_in, proportion_difference, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = 0) +
  facet_wrap(~race) +
  labs(
    x = "Opt-In Rate",
    y = "Proportion Error in Mean Wage and Salary Income",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())
  
```

Increasing opt in for small groups can increase their utility. 

### Percentiles

```{r}
#| label: percentiles
#| fig-height: 3

bind_cols(
  specs |>
    slice(rep(1:n(), each = 11)),
  results |>
    map("percentiles") |>
    map_dfr(.f = ~ .x |> filter(variable == "incwage"))
) |>
  mutate(proportion_difference = if_else(is.nan(proportion_difference), 0, proportion_difference)) |>
  mutate(opt_in = paste("Opt-In Rate:", opt_in)) |>
  ggplot(aes(p, proportion_difference, group = id)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~ opt_in, nrow = 1) +
  labs(
    x = "Salary and Income Percentile",
    y = "Proportion Error in Salary and Income Percentile"
  ) +
  scatter_grid() +
  theme(panel.spacing = ggplot2::unit(16L, "pt"))

bind_cols(
  specs |>
    slice(rep(1:n(), each = 11)),
  results |>
    map("percentiles") |>
    map_dfr(.f = ~ .x |> filter(variable == "ftotinc"))
) |>
  mutate(proportion_difference = if_else(is.nan(proportion_difference), 0, proportion_difference)) |>
  mutate(opt_in = paste("Opt-In Rate:", opt_in)) |>  
  ggplot(aes(p, proportion_difference, group = id)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~ opt_in, nrow = 1) +
  labs(
    x = "Family Income Percentile",
    y = "Proportion Error in Family Income Percentile"
  ) +
  scatter_grid() +
  theme(panel.spacing = ggplot2::unit(16L, "pt"))

bind_cols(
  specs |>
    slice(rep(1:n(), each = 11)),
  results |>
    map("percentiles") |>
    map_dfr(.f = ~ .x |> filter(variable == "inctot"))
) |>
  mutate(proportion_difference = if_else(is.nan(proportion_difference), 0, proportion_difference)) |>
  mutate(opt_in = paste("Opt-In Rate:", opt_in)) |>  
  ggplot(aes(p, proportion_difference, group = id)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~ opt_in, nrow = 1) +
  labs(
    x = "Total Income Percentile",
    y = "Proportion Error in Total Income Percentile"
  ) +
  scatter_grid() +
  theme(panel.spacing = ggplot2::unit(16L, "pt"))

bind_cols(
  specs |>
    slice(rep(1:n(), each = 11)),
  results |>
    map("percentiles") |>
    map_dfr(.f = ~ .x |> filter(variable == "incresid"))
) |>
  mutate(opt_in = paste("Opt-In Rate:", opt_in)) |>  
  mutate(proportion_difference = if_else(is.nan(proportion_difference), 0, proportion_difference)) |>
  ggplot(aes(p, proportion_difference, group = id)) +
  geom_line(alpha = 0.5) +
  facet_wrap(~ opt_in, nrow = 1) +
  labs(
    x = "Income Residual Percentile",
    y = "Proportion Error in Residual Income Percentile"
  ) +
  scatter_grid() +
  theme(panel.spacing = ggplot2::unit(16L, "pt"))

```

## Bivariate Utility

```{r}
#| label: correlation-difference
#| fig-height: 4

bind_cols(
  specs,
  correlation_difference_mae = results |>
    map_dbl(.f = ~ .x$cor_fit$correlation_difference_mae)
) |>
  ggplot(aes(opt_in, correlation_difference_mae, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  scale_y_continuous(limits = c(0, NA)) +
  labs(
    x = "Opt-In Rate",
    y = "Mean Absolute Error in Pairwise Correlation Coefficient",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

bind_cols(
  specs,
  correlation_difference_mae = results |>
    map_dbl(.f = ~ .x$cor_fit$correlation_difference_mae)
) |>
  group_by(opt_in, white_multiplier) |>
  summarize(correlation_difference_mae = median(correlation_difference_mae)) |>
  ungroup() |>
  ggplot(aes(opt_in, correlation_difference_mae, color = white_multiplier, group = white_multiplier)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Opt-In Rate",
    y = "Median Mean Absolute Error in Pairwise Correlation Coefficient",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

```

## Multivariate Utility

```{r}
#| label: kmarginals
#| fig-height: 4


k1_benchmark <- kmarginals(
  postsynth = holdout, 
  data = gsds, 
  k = 1
)

bind_cols(
  specs,
  k1_marginals = results |>
    map_dbl("k1_marginals")
) |>
  ggplot(aes(opt_in, k1_marginals, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = k1_benchmark, color = "red") +
  scale_y_continuous(limits = c(NA, 1000)) +
  labs(
    subtitle = "Score from Holdout Data in Red",
    x = "Opt-In Rate",
    y = "1-Marginal Score",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

k2_benchmark <- kmarginals(
  postsynth = holdout, 
  data = gsds, 
  k = 2
)

bind_cols(
  specs,
  k2_marginals = results |>
    map_dbl("k2_marginals")
) |>
  ggplot(aes(opt_in, k2_marginals, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = k2_benchmark, color = "red") +
  scale_y_continuous(limits = c(NA, 1000)) +  
  labs(
    subtitle = "Score from Holdout Data in Red",
    x = "Opt-In Rate",
    y = "2-Marginal Score",
    color = "White Multiplier"
  ) +  
  scatter_grid() +
  theme(legend.title = element_text())

k3_benchmark <- kmarginals(
  postsynth = holdout, 
  data = gsds, 
  k = 3
)

bind_cols(
  specs,
  k3_marginals = results |>
    map_dbl("k3_marginals")
) |>
  ggplot(aes(opt_in, k3_marginals, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = k3_benchmark, color = "red") +  
  scale_y_continuous(limits = c(NA, 1000)) +  
  labs(
    subtitle = "Score from Holdout Data in Red",
    x = "Opt-In Rate",
    y = "3-Marginal Score",
    color = "White Multiplier"
  ) +    
  scatter_grid() +
  theme(legend.title = element_text())

k1 <- bind_cols(
  specs,
  k1_marginals = results |>
    map_dbl("k1_marginals")
) |>
  ggplot(aes(opt_in, k1_marginals, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = k1_benchmark, color = "red") +
  scale_y_continuous(limits = c(NA, 1000)) +
  labs(
    x = "Opt-In Rate",
    y = "1-Marginal Score",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

k3 <- bind_cols(
  specs,
  k3_marginals = results |>
    map_dbl("k3_marginals")
) |>
  ggplot(aes(opt_in, k3_marginals, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = k3_benchmark, color = "red") +  
  scale_y_continuous(limits = c(NA, 1000)) +  
  labs(
    x = "Opt-In Rate",
    y = "3-Marginal Score",
    color = "White Multiplier"
  ) +    
  scatter_grid() +
  theme(legend.title = element_text())

k1 + k3

```

As `k` increases, the opt-in has a much bigger relative impact than the white multiplier. 

```{r}
#| label: discriminator-auc
#| fig-height: 3

bind_cols(
  specs,
  discriminator_auc = results |>
    map_dbl("discriminator_auc")
) |>
  ggplot(aes(opt_in, discriminator_auc, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    y = "AUC for Discrimnator Model",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

```

## Specific Utility

```{r}
#| label: regression-overlap 
#| fig-height: 4

regression_ci_overlap <- bind_cols(
  specs,
  discriminator_auc = results |>
    map_dfr(
      .f = ~ .x$regression_ci_overlap$ci_overlap |>
        summarize(
          ci_overlap = mean(overlap),
          abs_std_coef_diff = mean(abs(std_coef_diff)),
          sign_match = mean(sign_match),
          inference_match = mean(inference_match)
        )
    )
)

# CI difference
regression_ci_overlap |>
  ggplot(aes(opt_in, abs_std_coef_diff, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    y = "Absolute Standardized Coefficient Difference",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

# overlap
regression_ci_overlap |>
  ggplot(aes(opt_in, ci_overlap, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    y = "Regression Confidence Interval Overlap",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

# sign match
regression_ci_overlap |>
  ggplot(aes(opt_in, sign_match, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

# significance match
regression_ci_overlap |>
  ggplot(aes(opt_in, inference_match, color = white_multiplier)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Opt-In Rate",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

```

## Disclosure Metrics

### L-Diversity

```{r}
bind_cols(
  specs,
  discriminator_auc = results |>
    map_dfr("ldiveristy")
) |>
  group_by(opt_in, white_multiplier) |>
  summarize(
    max(ldiveristy_age),
    max(ldiveristy_educd),
    max(ldiveristy_ftotinc),
    max(ldiveristy_inctot),
    max(ldiveristy_incwage)
  )

```

### Attribute Inference Test

```{r}
#| label: attribute-inference

holdout_rmse <- prediction_test(
  postsynth = holdout, 
  data = gsds, 
  formula = incwelfr ~ .
)

bind_cols(
  specs,
  incwelfr_rmse = results |>
    map_dbl("incwelfr_rmse")
) |>
  ggplot(aes(opt_in, incwelfr_rmse)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = holdout_rmse, color = "red") +
  labs(
    x = "Opt-In Rate",
    y = "Welfare Income RMSE"
  )
  scatter_grid()

```

### Membership Inference Test

```{r}
#| label: membership-inference
#| fig-height: 3.2

bind_cols(
  specs,
  membership_inference_auc = results |>
    map("membership_inference_test") |>
    map_dbl("auc")
) |>
  ggplot(aes(opt_in, membership_inference_auc, color = white_multiplier)) +
  geom_hline(aes(yintercept = 0.5), color = "red") +
  geom_point(alpha = 0.5) +
  labs(
    x = "Opt-In Rate",
    y = "Membership Inference Test AUC",
    color = "White Multiplier"
  ) +
  scatter_grid() +
  theme(legend.title = element_text())

results |>
    map("membership_inference_test") |>
    map("conf_mat")

```



